---
title: 'Tidy analysis #1 : Correlation and regression fundamentals with tidy data principles'
author: "Jacob Lowell"
date: "4/23/2020"
output: 
   html_document:
       theme: flatly
       toc: true
       toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , message = FALSE , warning =  FALSE)

library(knitr)
options(digits=3)

```

I am working through the first post on the tidymodels site, located at : https://www.tidymodels.org/learn/statistics/tidy-analysis/

# Section 1

## load data

```{r}
library(tidymodels)
library(tidyverse)
library(knitr)
library(gt)
options(scipen=999)

data(Orange)

Orange <- as_tibble(Orange)

cor(Orange$age, Orange$circumference)


ggplot(Orange, aes(age, circumference, color = Tree)) +
  geom_line()+ labs(title =  "Orange circumference by tree")
```



## test for correlation within each tree

```{r}
Orange %>% 
  group_by(Tree) %>%
  summarize(correlation = cor(age, circumference)) %>%  gt()

```


## hypothesis test circumference vs age

```{r}
ct <- cor.test(Orange$age, Orange$circumference)
tidy(ct ) %>%  gt()

tidy(ct )
```



## Tidy hypothesis test results with tidy() function

```{r}

```



# Section 2

## Nested data  - nest-map-unnest


```{r}
library(tidyr)

nested <- 
  Orange %>% 
  nest(data = c(age, circumference))
```


### correlation test for each tibble

```{r}
nested %>% 
  mutate(test = map(data, ~ cor.test(.x$age, .x$circumference)))
```

### Tidy each object

```{r}

nested %>% 
  mutate(
    test = map(data, ~ cor.test(.x$age, .x$circumference)), # S3 list-col
    tidied = map(test, tidy)
  ) 

```


### Unnest the tidied data

```{r}
Orange %>% 
  nest(data = c(age, circumference)) %>% 
  mutate(
    test = map(data, ~ cor.test(.x$age, .x$circumference)), # S3 list-col
    tidied = map(test, tidy)
  ) %>% 
  unnest(cols = tidied) %>% 
  select(-data, -test) %>% gt()
```



# Section 3 Regression Models

## More Useful when applied to Regression

```{r}
lm_fit <- lm(age ~ circumference, data = Orange)
tidy(lm_fit)

```

## Multiple regressions with nest -> map -> unnest

```{r}
Orange %>%
  nest(data = c(-Tree)) %>% 
  mutate(
    fit = map(data, ~ lm(age ~ circumference, data = .x)),
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied) %>% 
  select(-data, -fit) %>%  gt()
```




## MT cars to nest automatic vs manual transmition  


```{r}
data(mtcars)
mtcars <- as_tibble(mtcars)


mtcars %>%
  nest(data = c(-am)) %>% 
  mutate(
    fit = map(data, ~ lm(wt ~ mpg + qsec + gear, data = .x)),  # S3 list-col
    tidied = map(fit, tidy)
  ) %>% 
  unnest(tidied) %>% 
  select(-data, -fit) %>% gt()

```


### store tidy, augment and glance all at once! 

- Glance shows model performance 
- Tidy show coefficients
- Augmet shows observation level fit statistics


```{r}
regressions <- 
  mtcars %>%
  nest(data = c(-am)) %>% 
  mutate(
    fit = map(data, ~ lm(wt ~ mpg + qsec + gear, data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    augmented = map(fit, augment)
  )

regressions %>% 
  select(tidied) %>% 
  unnest(tidied) %>%  gt()

regressions %>% 
  select(augmented) %>% 
  unnest(augmented) %>%  gt()


regressions %>% 
  select(glanced) %>% 
  unnest(glanced) %>%  gt()

```


