---
title: 'Tidymodels #2 - K Means'
author: "Jacob Lowell"
date: "4/23/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , message = FALSE , warning =  FALSE)

library(knitr)
options(digits=3)

```

I am working through the first post on the tidymodels site, located at : https://www.tidymodels.org/learn/statistics/k-means/

# Section 1

## Generate data

```{r}
library(tidymodels)
library(tidyverse)


library(tidymodels)
library(tidyr)

set.seed(27)

centers <- tibble(
  cluster = factor(1:3), 
  num_points = c(100, 150, 50),  # number points in each cluster
  x1 = c(5, 0, -3),              # x1 coordinate of cluster center
  x2 = c(-1, 1, -2)              # x2 coordinate of cluster center
)

labelled_points <- 
  centers %>%
  mutate(
    x1 = map2(num_points, x1, rnorm),
    x2 = map2(num_points, x2, rnorm)
  ) %>% 
  select(-num_points) %>% 
  unnest(cols = c(x1, x2))

ggplot(labelled_points, aes(x1, x2, color = cluster)) +
  geom_point(alpha = 0.3)

```



## Cluster


```{r}
points <- 
  labelled_points %>% 
  select(-cluster)

kclust <- kmeans(points, centers = 3)
kclust

summary(kclust)


```


## Check out augment, tidy and glance

* cluster (300 values) contains information about each point
* centers, withinss, and size (3 values) contain information about each cluster
* totss, tot.withinss, betweenss, and iter (1 value) contain information about the full clustering


This applies kluster to the raw data

```{r}

# observation level

augment(kclust , points)


# stats for each cluster
tidy(kclust)


# summary statistics 
glance(kclust)

```


## Try clusters of multiple sizes, see what is best

```{r}
kclusts <- 
  tibble(k = 1:9) %>%
  mutate(
    kclust = map(k, ~kmeans(points, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, points)
  )

kclusts
```



##  three different data sets for clusterin insight

```{r}
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))
```






Now we can plot the original points using the data from augment(), with each point colored according to the predicted cluster.
```{r}


p1 <- 
  ggplot(assignments, aes(x = x1, y = x2)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1

```

### add an X for the center of each cluster

```{r}

p2 <- p1 + geom_point(data = clusters, size = 10, shape = "x")
p2
```




###  Summary statistics across k with glance()

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()

```

This represents the variance within the clusters. It decreases as k increases, but notice a bend (or “elbow”) around k = 3. This bend indicates that additional clusters beyond the third have little value.